Computing Entropies With Nested Sampling

### 2017-07-30 paper/ms.tex
Putting q(x) in, as Skilling would like to see

If an unknown quantity $x$ has a discrete probability distribution $p(x)$,
the Shannon entropy \citep{shannon, cover2012elements} is defined as
\begin{align}
H(x) &= -\sum_{x} p(x) \log p(x)
\end{align}
where the sum is over all of the possible values of $x$ under consideration.
The entropy quantifies the degree to which the issue
``what is the value of $x$, precisely?'' remains unresolved
\citep{knuth_questions}. For example,
if there is only one possible value which has probability 1, $H(x)$ is
zero. Conventionally, $0 \times \log 0$ is defined to be equal
to $\lim_{x \to 0^+} (x\log x)$, which is zero; i.e., `possibilities' with
probability zero do not contribute to the sum.
If there are $N$ possibilities with equal probabilities $1/N$ each,
then the entropy is $H(x) = \log(N)$.
If the space of possible $x$-values is continuous so that $p(x)$ is a
probability density function, the differential entropy
\begin{align}
H(x) &= -\int p(x) \log p(x) \, dx
\end{align}
quantifies uncertainty by generalising the log-volume
of the plausible region, when volume is defined with respect to $dx$.

Entropies, and related quantities, tend to be analytically available only for
a few families of probability distributions. On the numerical side, if
$\log p(x)$ can be evaluated for any $x$, then simple Monte Carlo will suffice
for approximating $H(x)$. On the other hand, if $p(x)$ can only be
{\em sampled} but not evaluated (for example, if $p(x)$ is a marginal
distribution), then this will not work. Kernel density estimation or the like
\citep[e.g.][]{JMLR:v15:szabo14a}
may be effective in this case, but is unlikely to generalise well to high
dimensions.

This paper introduces a computational approach to evaluating the entropy 
$H(x)$ of probability distributions that can only be sampled
using Markov Chain Monte Carlo (MCMC).

>>>>

If an unknown quantity $x$ has a discrete probability distribution $p(x)$,
the Shannon entropy \citep{shannon, cover2012elements} is defined as
\begin{align}
H(x) &= -\sum_{x} p(x) \log p(x)
\end{align}
where the sum is over all of the possible values of $x$ under consideration.
The entropy quantifies the degree to which the issue
``what is the value of $x$, precisely?'' remains unresolved
\citep{knuth_questions}. More generally, the `Kullback-Leibler divergence'
quantifies the degree to which a probability distribution $p(x)$ departs
from a `base' or `prior' probability distribution $q(x)$:
\begin{align}
D_{\rm KL}\big(p \,||\, q\big) &=
    \sum_{x} p(x) \log\left[\frac{p(x)}{q(x)}\right]
\end{align}
\citep{knuth2012foundations, caticha2006updating}.
If $q(x)$ is taken to be uniform, this is equivalent (up to an additive
constant and a change of sign) to the Shannon entropy $H(x)$.
With the negative sign reinstated, Kullback-Leibler divergence is sometimes
called relative entropy.

The Shannon entropy can be interpreted straightforwardly as a measure of
uncertainty. For example,
if there is only one possible value which has probability 1, $H(x)$ is
zero. Conventionally, $0 \times \log 0$ is defined to be equal
to $\lim_{x \to 0^+} (x\log x)$, which is zero; i.e., `possibilities' with
probability zero do not contribute to the sum.
If there are $N$ possibilities with equal probabilities $1/N$ each,
then the entropy is $H(x) = \log(N)$.
If the space of possible $x$-values is continuous so that $p(x)$ is a
probability density function, the differential entropy
\begin{align}
H(x) &= -\int p(x) \log p(x) \, dx
\end{align}
quantifies uncertainty by generalising the log-volume
of the plausible region, where volume is defined with respect to $dx$.
Crucially, in continuous spaces, the differential entropy is not
invariant under changes of coordinates, so any value given for a
differential entropy must be understood with respect to the coordinate
system that was used.

Entropies, relative entropies, and related quantities,
tend to be analytically available only for
a few families of probability distributions. On the numerical side, if
$\log p(x)$ can be evaluated for any $x$, then simple Monte Carlo will suffice
for approximating $H(x)$. On the other hand, if $p(x)$ can only be
{\em sampled} but not evaluated (for example, if $p(x)$ is a marginal
distribution), then this will not work. Kernel density estimation or the like
\citep[e.g.][]{JMLR:v15:szabo14a}
may be effective in this case, but is unlikely to generalise well to high
dimensions.

This paper introduces a computational approach to evaluating the Shannon
entropy $H(x)$ of probability distributions that can only be sampled
using Markov Chain Monte Carlo (MCMC). With minor modifications,
Kullback-Leibler divergences may also be obtained.
